---
title: "Machine Learning | Course Project"
author: "Taurã Figueiredo"
date: "Saturday, August 23, 2014"
output: html_document
---
## Machine learning algorithm

Download files provided in the assignment and load into R   
```{r eval=FALSE, warning=FALSE}
download.file(url = "https://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv", destfile = "pml-training.csv")
download.file(url = "https://d396qusza40orc.cloudfront.net/predmachlearn/pml-testing.csv", destfile = "pml-testing.csv")
```
```{r warning=FALSE}
pmlTraining <- read.csv("pml-training.csv")
pmlTesting <- read.csv("pml-testing.csv")
```

The first thing we should do is cleanup the data to keep only the relevant information.   
So, to do that, I create a function that remove every collumn that have more than 50% of it rows with empty data, NA data and not a number data. Also, before return the new dataframe, I remove the first 4 collumns that are timestamp or not relevant info to predict.

```{r warning=FALSE}
# Remove empty, NA and not useful variables
ProcessData <- function (dataframe){
  col <- vector(mode = "numeric")
  count <- 0
  for(i in 1:ncol(dataframe))
  {
    if(colnames(dataframe)[i]=='classe'){next;}
    
    total <- length(dataframe[,i]);
    
    NAs <- length(dataframe[is.na(dataframe[,i]),i]);
    empty <- length(dataframe[dataframe[,i]=='',i])
    notNumber <- length(dataframe[!is.numeric(dataframe[,i]),i]);
    
    if(round(NAs/total)==1 | round(empty/total)==1 | round(notNumber/total)==1){
      col <- rbind(col, i)
      count <- count+1;
    }
  }
  
  process <- dataframe[,as.numeric(col)*(-1)]
  for(i in 1:(ncol(process)-1)){ process[,i] <- as.numeric(process[,i]); }
  process[,c(1,2,3,4)*(-1)]
}
```

So, now that I have a function to pre process the data, I need to pre process the Training data.   
```{r warning=FALSE}
cleanedData <- ProcessData(pmlTraining)
```

With all the data processed, it's time to start to train our model.   
   
First, load the packages that I will use (caret and randomForest).   
Then, split the training dataset into two new datasets: trainSet and validationSet.
```{r warning=FALSE}
library(caret)
library(randomForest)

inTrain <- createDataPartition(y = cleanedData$classe, p = 0.7, list = FALSE)

trainSet <- cleanedData[inTrain,]
validationSet <- cleanedData[-inTrain,]
```

Now it's time to train the model using randomForest.
```{r warning=FALSE, cache=TRUE}
fit <- randomForest(classe ~ ., data = trainSet)
```

With the model trained, let's test with our validationSet.
```{r warning=FALSE}
confusionMatrix(predict(fit, validationSet[,-53]), validationSet$classe)
```

Ok. Now I got a very accurate model.   
Let's predict the pmlTesting.   

Before just predict, we need to apply the same procedure done to our training set.   
So, I just used my ProcessData function.   
Then, it was just predict.
```{r warning=FALSE}
testSet <- ProcessData(pmlTesting)
predict(fit, testSet)
```